---
title: "Question 2"
author: "Joseph Froelicher"
date: "4/9/2021"
output: pdf_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE, include = TRUE)

```

## Part A
```{r a}
skincancer = read.csv("skincancer.csv", header = TRUE)

for (i in 1:dim(skincancer)[1]){
  if (skincancer$city[i] == 'Dallas') {
    skincancer$city[i] = 1
  } else {
    skincancer$city[i] = 0
  }
}

model1 = glm(cases ~ city + offset(log(py1000)) + age.group, family = "poisson", data = skincancer)
table = summary(model1)$coefficients

rownames(table) = c(
  'Minneappolis (B0)', 'Dallas (B1)', '25-34', '35-44', '45-54', '55-64', '65-74', '75-84', '85+'
)

colnames(table) = c('Estimate', 'Std. Error', 'Z-value', 'p-value')
table

```

## Part B
* $H_0$: $e^{\beta_{Dallas}} = 2$ which means $\beta_{Dallas} = \log(2)$
* $H_A$: $e^{\beta_{Dallas}} \ne 2$
* Test statistic: 

The Wald test statistic is given by:
$$
Z^2 = \frac{(\hat{\beta}_{Dallas} - \beta_{H_0})^2}{Var(\hat{\beta}_{Dallas})} \sim \chi^2_1
$$

$$
Z^2 = \frac{( 0.787 - \log(2))^2}{0.0514^2} \sim \chi^2_1
$$

```{r b}
Z = (coef(model1)[2] - log(2))/sqrt(vcov(model1)[2,2])
p = pchisq(Z^2, df = 1, lower.tail = FALSE)

```

We conclude that the there is not evidence to suggest that the age-adjusted non-melanoma skin cancer rates in Dallas for women was *not* double that of women in Minneapolis (p = `r round(p, 3)`).

## Part C
* $H_0$: $\beta_{25 to 34} = 0$ (15 to 24 is the reference group)
* $H_A$: $\beta_{25 to 34} \ne 0$
* Test statistic and Null distribution

The test statistic reported by R is the Wald test statistic, Z, which follows a standard normal distribution under the null hypothesis.

$$Z = 4.77 \sim N(0, 1)$$
We conclude that there is significant evidence to suggest that rates in the 25 to 34 age group are not equal to those in the 15 to 24 age group (p < 0.001). 

## Part D
```{r d}
r = exp(sum(c(1, 0, 0, 0, 1, 0, 0, 0, 0) * coef(model1)))

```
The estimated rate of non-melanoma skin cancer in the 45 to 54 year age group in women is `r round(r, 3)`.
