---
title: "Question 1"
author: "Joseph Froelicher"
date: "4/9/2021"
output: pdf_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE, include = TRUE)

```

## Part A
```{r a}
skincancer = read.csv("skincancer.csv", header = TRUE)

for (i in 1:dim(skincancer)[1]){
  if (skincancer$city[i] == 'Dallas') {
    skincancer$city[i] = 1
  } else {
    skincancer$city[i] = 0
  }
}

model0 = glm(cases ~ city + offset(log(py1000)), family = "poisson", data = skincancer)
table = summary(model0)$coefficients
rownames(table) = c('Minneappolis (B0)', 'Dallas (B1)')
colnames(table) = c('Estimate', 'Std. Error', 'Z-value', 'p-value')
ci = exp(confint.default(model0))
table

```

## Part B
* $\beta_0$: the log rate of non-melanoma skin cancer for women in Minneapolis.
* $\beta_1$: the log rate ratio of non-melanoma skin cancer for women in Dallas vs. women of Minneapolis.

* $e^{\beta_0}$: The rate of non-melanoma skin cancer for women in Minneapolis is `r exp(table[1, 1])` (95% CI: [`r ci[1, 1]`, `r ci[1, 2]`]) cases per 1000 person years.
* $e^{\beta_1}$: The rate of non-melanoma skin cancer for women in Dallas in cases per 1000 person years is `r exp(table[2, 1])` (95% CI: [`r ci[2, 1]`, `r ci[2, 2]`]) times the rate of non-melanoma skin cancer for women of Minneapolis.



