---
title: "Question 3"
author: "Joseph Froelicher"
date: "2/26/2021"
output: pdf_document
---

```{r setup, echo = FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part A
The odds of the adolescent having obtained condoms increased by a a factor of 4.04 for those adolescents who took part in an HIV educational program. The corresponding confidence interval was [1.17, 13.9], which does not contain 1, hence, there was a significant increase in the odds of an adolescent having obtained condoms if they took part in an HIV educational program.

# Part B
```{r b, echo = FALSE}
# part B
b1 = log(4.04)
b2 = log(1.38)
b3 = log(5.82)
b4 = log(3.22)

```
$\hat{\beta_1}=$ `r b1`
$\hat{\beta_2}=$ `r b2`
$\hat{\beta_3}=$ `r b3`
$\hat{\beta_4}=$ `r b4`

# Part C
We would need to know the probability of buying condoms when group, gender, SES and pifetime partners are all equal to 0.

# Part D
```{r d, echo = FALSE}
# part D
l = log(1.17)
u = log(13.9)

# Does b1 = (l + u) / 2?
m = (l + u) / 2 # Close enough

a = 0.05
z = qnorm(1 - (a / 2))

se1 = (m - l) / z
se2 = (u - m) / z  # double check

# reconstruct to triple check
ci = c(exp(b1 - (z * se1)), exp(b1 + (z * se2))) # noice

```
Based on the corresponding Wald 95% confidence interval for the log odds ratio, the standard error for the `group` effect is `r se1`.

# Part E
```{r e, echo = FALSE}
# part E
l2 = log(1.23)
u2 = log(12.88)

# Does b2 = (l2 + u2) / 2?
m2 = (l2 + u2) / 2 # Close enough

se12 = (m2 - l2) / z
se22 = (u2 - m2) / z  # double check

# reconstruct to triple check
ci2 = c(exp(b2 - (z * se12)), exp(b2 + (z * se22))) # not noice
ci3 = c(exp(exp(b2) - (z * se12)), exp(exp(b2) + (z * se22))) # ahhh much better

```
Based on our confidence interval, the standard error for gender should be `r se12`. however, if we try to reconstruct that confidence interval using the provided odds ratio, we get [`r ci2[1]`, `r ci2[2]`], which doesn't match our provided confidence interval. However, if we use the provided 3.97 in the question as the odds ratio to reconstruct the confidence interval, we get [`r ci3[1]`, `r ci3[2]`], which matches the provided confidence interval. Hence, the actual odds ratio should read `r exp(exp(b2))`.
\newpage

# Appendix
```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```