---
author: "Joseph Froelicher"
title: "Question 2"
date: "2/26/2021"
output: pdf_document
---

```{r setup, echo = FALSE, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
data = read.table("copd.txt", header = T)

library(aod)

```

# Part A
```{r a, echo = FALSE, include = FALSE}
# part A
model1 = glm(copd ~ age + gender + smoker + BMI + BMIsquared, data = data, family = binomial)
summary(model1)
t_w = summary(model1)$coefficients[6, 3]
p_w = 2 * (1 - pnorm(t_w))

```
COPD is significantly associated with BMI squared, based on a wald test (T(wald) = `r t_w`, p = `r p_w`).

# Part B
```{r b, echo = FALSE, include = FALSE}
# part B
model0 = glm(copd ~ age + gender + smoker + BMI, data = data, family = binomial)
t_lr = (-2 * log(exp(logLik(model0)) / exp(logLik(model1))))[1]
df = 1
p_lr = pchisq(t_lr, df, lower.tail = F)

```
COPD is significantly associated with BMI squared, based on a wald test (T(likelihood ratio) = `r t_lr`, p = `r p_lr`).

# Part C
Yes, both the wald test and likelihood ration test indicated that there is evidence that COPD has a quadratic relationship with BMI ($\alpha$ = 0.05).

# Part D
Since we are using a higher order term for BMI, in order to avoid high correlation between BMI and its higher order term, BMI squared, we should center the variable.
\newpage

# Appendix
```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```