---
author: "Joseph Froelicher"
title:  "Question 1"
date: "February 26, 2020"
output: pdf_document
---
\usepackage{bm}

# Part A
$$ln\left(\frac{p_i}{1-p_i}\right) = \textbf{X}_j\mathbf{\beta}$$
$$\hat{\beta_0} = ln\left(\frac{p(y=1|x_j=0)}{1-p(y=1|x_j=0)}\right) = ln\left(\frac{(619/2416)}{1-(619/2416)}\right)=-1.065769$$
$$\hat{\beta_1}=(\hat{\beta_0}+\hat{\beta_1})-\hat{\beta_0}=ln\left(\frac{p(y=1|x_j=1)}{1-p(y=1|x_j=1)}\right)-\hat{\beta_0} = ln\left(\frac{(355/771)}{1-(355/771)}\right)-(-1.065769)=0.9072015$$
$$\hat{\beta_2}=(\hat{\beta_0}+\hat{\beta_2})-\hat{\beta_0}=ln\left(\frac{p(y=1|x_j=2)}{1-p(y=1|x_j=2)}\right)-\hat{\beta_0} = ln\left(\frac{(162/731)}{1-(162/731)}\right)-(-1.065769)=-0.1905151$$

# Part B
$$L(Y_i|p)=\prod_{i=1}^N{n_i\choose Y_i}p^{Y_i}(1-p)^{n_i-Y_i}$$
$$L(Y_i|p)=\prod_{i=1}^N{n_i\choose Y_i}\left(\frac{e^{\textbf{X}\beta}}{1+e^{\textbf{X}\beta}}\right)^{Y_i}\left(1-\frac{e^{\textbf{X}\beta}}{1+e^{\textbf{X}\beta}}\right)^{n_i-Y_i}$$

$$L(Y_i|p)=\left(\frac{e^{\textbf{X}\beta}}{1+e^{\textbf{X}\beta}}\right)^{\sum_{i=1}^nY_i}\left(1-\frac{e^{\textbf{X}\beta}}{1+e^{\textbf{X}\beta}}\right)^{n-\sum_{i=1}^nY_i}$$
$$ln(L(Y_i|p))=ln\left(\left(\frac{e^{\textbf{X}\beta}}{1+e^{\textbf{X}\beta}}\right)^{\sum_{i=1}^nY_i}\left(1-\frac{e^{\textbf{X}\beta}}{1+e^{\textbf{X}\beta}}\right)^{n-\sum_{i=1}^nY_i}\right)$$
$$l(Y_i|p)=\sum_{i=1}^n\left(ln{n\choose Y_i}+Y_i\:ln\left(\frac{e^{\textbf{X}\beta}}{1+e^{\textbf{X}\beta}}\right)+(n-Y_i)\:ln\left(1-\frac{e^{\textbf{X}\beta}}{1+e^{\textbf{X}\beta}}\right)\right)$$
$$l(Y_i|p)=\left[ln{n_i\choose Y_i}+Y_i\:ln\left(\frac{e^{\beta_0}}{1+e^{\beta_0}}\right)+(n_i-Y_i)\:ln\left(1-\frac{e^{\beta_0}}{1+e^{\beta_0}}\right)\right]$$
$$\;\;\;+\;\left[ln{n_i\choose Y_i}+Y_i\:ln\left(\frac{e^{\beta_0+X_1\beta_1}}{1+e^{\beta_0+X_1\beta_1}}\right)+(n_i-Y_i)\:ln\left(1-\frac{e^{\beta_0+X_1\beta_1}}{1+e^{\beta_0+X_1\beta_1}}\right)\right]$$
$$\;\;\;+\;\left[ln{n_i\choose Y_i}+Y_i\:ln\left(\frac{e^{\beta_0+X_2\beta_2}}{1+e^{\beta_0+X_2\beta_2}}\right)+(n_i-Y_i)\:ln\left(1-\frac{e^{\beta_0+X_2\beta_2}}{1+e^{\beta_0+X_2\beta_2}}\right)\right]$$

```{r b}
data = data.frame(
  "strikes" = c("1", "2", "3"),
  "misconduct" = c(619, 355, 162),
  "no_misconduct" = c(1797, 416, 569)
)

# by hand
p1 = data[1, 2] / (data[1, 2] + data[1, 3])
p2 = data[2, 2] / (data[2, 2] + data[2, 3])
p3 = data[3, 2] / (data[3, 2] + data[3, 3])

n1 = data[1, 2] + data[1, 3]
n2 = data[2, 2] + data[2, 3]
n3 = data[3, 2] + data[3, 3]

c1 = data[1, 2]
c2 = data[2, 2]
c3 = data[3, 2]

w1 = data[1, 3]
w2 = data[2, 3]
w3 = data[3, 3]
  
ll_m1 = lchoose(n1, c1) + c1 * log(p1) + w1 * log(1 - p1)
ll_m2 = lchoose(n2, c2) + c2 * log(p2) + w2 * log(1 - p2)
ll_m3 = lchoose(n3, c3) + c3 * log(p3) + w3 * log(1 - p3)

model1 = glm(
  cbind(misconduct, no_misconduct) ~ strikes, data, family = binomial
)

like1 = logLik(model1)
```
Using the `logLik()` function in R, we get the value of `r like1` for the likelihood of model 1.

# Part C
$$l(Y_i|p)=\sum_{i=1}^n\left(Y_i\:ln(p_i)+(n-Y_i)\:ln(1-p_i)\right)$$
$$l(Y_i, X_{ij}|\beta)=\sum_{i=1}^n\left(ln{n\choose Y_i}+Y_i\:ln(\frac{e^{\beta_0}}{1-e^{\beta_0}})+(1-Y_i)\:ln(\frac{e^{\beta_0}}{1-e^{\beta_0}})\right)$$
$$l(Y_i, X_{ij}|\beta)=\sum_{i=1}^n\left(Y_i\:ln(\frac{e^{\beta_0}}{1-e^{\beta_0}})+(1-Y_i)\:ln(\frac{e^{\beta_0}}{1-e^{\beta_0}})\right)$$

```{r c}
model0 = glm(
  cbind(misconduct, no_misconduct) ~ 1, data, family = binomial
)

like0 = logLik(model0)

```

Using the `logLik()` function in R, we get the value of `r like0` for the likelihood of model 0.

# Part D
$$T_{LR} = -2ln\left[\frac{L(p_{H_0})}{L(p_{MLE})}\right]$$
$$T_{LR} = -2ln\left[\frac{e^{-76.41084}}{e^{-10.86998}}\right]$$
$$X_{2, 1-0.95}^2 = 5.991$$
```{r d}
pchisq(131.0817, 3, lower.tail = F)

```
$$P\{(T_{LR} = 131.0817) > 5.991\} < 0.001$$
$$(T_{LR} = 131.0817) \sim X_{3, 0.95}^2$$

# Part E
```{r}
data_numeric = data.frame(
  "strikes" = c(1, 2, 3),
  "misconduct" = c(619, 355, 162),
  "no_misconduct" = c(1797, 416, 569)
)

model2 = glm(
  cbind(misconduct, no_misconduct) ~ strikes, data_numeric, family = binomial
)

summary(model2)
like2 = logLik(model2)

```

The predicted probability of a misconduct violation during the first year in prison for a prisoner with 1 strike is `r p1`. The predicted probability of a misconduct violation during the first year in prison for a prisoner with 3 strikes is `r p3`.

# Part F
```{r}
p3 / p1
```
$$CI = [e^{\hat{\beta}-Z_{1-\frac{\alpha}{2}}\hat{SE}(\hat{\beta})}, e^{\hat{\beta}+Z_{1-\frac{\alpha}{2}}\hat{SE}(\hat{\beta})}]$$

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```