---
author: "Joseph Froelicher"
title: "Question 1"
date: "3/12/2021"
output: pdf_document
---

```{r setup, echo = FALSE, include = FALSE}
n = 20
k = rep(1 / 100, n * 3)

data = data.frame(
  "hours" = c(8, 12, 16),
  "pos" = c(5, 10, 19)
)

data.glm = data.frame(
  "hours" = c(rep(data$hours[1], n), rep(data$hours[2], n), rep(data$hours[3], n)),
  "pos" = c(
    rep(1, data$pos[1]),
    rep(0, n - data$pos[1]),
    rep(1, data$pos[2]),
    rep(0, n - data$pos[2]),
    rep(1, data$pos[3]),
    rep(0, n - data$pos[3])
  )
)

```

## Part A
$$f(y|p) = p^y(1-p)^{1-y},\;\;\;\;\;p = 1-e^{-k2^{t/\theta}}$$
$$f(y|\theta) = (1 -e^{-k2^{t/\theta}})^y(-e^{-k2^{t/\theta}})^{1-y}$$
$$L(\theta|\mathbf{y}) = \prod_{i=1}^n(1 -e^{-k2^{t/\theta}})^y(-e^{-k2^{t/\theta}})^{1-y}$$
$$L(\theta|\mathbf{y}) = (1 -e^{-k2^{t/\theta}})^{\sum_{i=1}^ny_i}(-e^{-k2^{t/\theta}})^{n-\sum_{i=1}^ny_i}$$
$$l(\theta|\mathbf{y}) = \sum_{i=1}^ny_i\;ln(1 -e^{-k2^{t/\theta}}) + (n-\sum_{i=1}^ny_i)\;ln(-e^{-k2^{t/\theta}})$$
$$l(\theta|\mathbf{y}) = \sum_{i=1}^ny_i\;(ln(1)-(-k2^{t/\theta})) + (n-\sum_{i=1}^ny_i)\;(-k2^{t/\theta})$$
$$l(\theta|\mathbf{y}) = \sum_{i=1}^ny_i\;(k2^{t/\theta}) + (n-\sum_{i=1}^ny_i)\;(-k2^{t/\theta})$$
$$l(\theta|\mathbf{y},t=\{8, 12, 16\}) = \sum_{i=1}^ny_i\;(k2^{8/\theta}) + (n-\sum_{i=1}^ny_i)\;(-k2^{8/\theta})$$
$$\;\;\;\;\;\;\;\;\;\;\;+\;\sum_{i=1}^ny_i\;(k2^{12/\theta}) + (n-\sum_{i=1}^ny_i)\;(-k2^{12/\theta})$$
$$\;\;\;\;\;\;\;\;\;\;\;+\;\sum_{i=1}^ny_i\;(k2^{16/\theta}) + (n-\sum_{i=1}^ny_i)\;(-k2^{16/\theta})$$

## Part B

$$g(e[Y_i|X_i]) = \mathbf{X}_i^T\beta,\;\;\;\;g(p_i) = ln(-ln(1 - p_i)),\;\;\;\;\mathbf{X}_i^T\beta=\beta_1t_i + ln(k)$$
From our model:
$$p_i = 1-e^{-e^{\beta_1t_i + ln(k)}}$$
$$p_i = 1-e^{-ke^{\beta_1t_i}}$$

Given in the problem:
$$p_i = 1-e^{-k2^{t_i/\theta}}$$
We know that:
$$\theta = f(\beta_1)$$
And thus by the invariance property of MLE's:
$$\hat{\theta} = f(\hat{\beta_1})$$

Therefore, Set them equal and solve:
$$1-e^{-ke^{\beta_1t_i}} = 1-e^{-k2^{t_i/\theta}}$$
$$-ke^{\beta_1t_i} = -k2^{t_i/\theta}$$

$$e^{\beta_1t_i} = 2^{t_i/\theta}$$
$$\beta_1t_i = \frac{t_i}{\theta}ln(2)$$
$$\beta_1 = \theta\;ln(2)$$
$$\hat{\theta} = \frac{ln(2)}{\hat{\beta_1}}$$
Q.E.D.

## Part C

```{r c, echo = TRUE, include = TRUE}
# Part C
model1 = glm(
  pos ~ hours + offset(log(k)) - 1, family = binomial(link = "cloglog"), data = data.glm
)
summary(model1)

```

## Part D

```{r d, echo = FALSE, include = FALSE}
# Part D
b1 = summary(model1)$coefficients[1, 1]
se = summary(model1)$coefficients[1, 2]
alpha = 0.05
z = qnorm(1 - (alpha / 2))
b1_ci = c(b1 - (z * se), b1 + (z * se))
theta = log(2) / b1
theta_ci = c(log(2) / b1_ci[1], log(2) / b1_ci[2])

```

$\hat{\theta}_{MLE}$ = `r theta` [`r theta_ci[2]`, `r theta_ci[1]`]

## Part E

```{r e, echo = FALSE, include = FALSE}
# part E
model0 = glm(
  pos ~ offset(log(k)) - 1, family = binomial(link = "cloglog"), data = data.glm
)
summary(model0)
anova(model1, model0, test = "LRT")

```

We reject the null hypothesis that the full model containing the time group variable is not a better fit that the saturated model containing no time group variable ($p < 0.001$).

## Appendix
```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
