---
author: "Joseph Froelicher"
title: "Question 2"
date: "3/12/2021"
output: pdf_document
---

```{r setup, echo = FALSE, include = FALSE}
library(tidyverse)

```

## Part A

```{r a, echo = FALSE, include = FALSE}
# Part A
n = 100000
s = 1
pde = 0.1
pdu = 0.05
pe = 0.3

b0 = log(pdu / (1 - pdu))
b1 = log(pde / (1 - pde) / (pdu / (1 - pdu)))

```

We need two parameters to simulate a logistic regression to model the outcome using the only the exposure, and those are $\beta_0$ and $\beta_1$. In order to calculate $\beta_0$ and $\beta_1$, we need the probability of disease, given exposure $P(Y = 1|X = 1)$, and the probability of disease, given no exposure $P(Y = 1|X = 0)$.
\newline
$$\beta_0 = ln\left(\frac{P(Y = 1|X = 0)}{1 - P(Y = 1|X = 0)}\right)$$
$$\beta_1 = ln\left(\frac{\left[\frac{P(Y = 1|X = 1)}{1 - P(Y = 1|X = 1)}\right]}{\left[\frac{P(Y = 1|X = 0)}{1 - P(Y = 1|X = 0)}\right]}\right)$$
$\beta_0$ = `r b0`,
$\beta_1$ = `r b1`

## Part B

```{r b, echo = TRUE, include = TRUE}
# Part B
set.seed(8675309)

x = rbinom(n, s, pe)
pee = plogis(b0 + b1 * x)
y = rbinom(n, s, pee)

data = data.frame(
  "id" = 1:n,
  "outcome" = y,
  "exposure" = x
)

# check population proportions
sume = 0
counte = 0
sumu = 0
countu = 0

for (i in 1:dim(data)[1]) {
  if (data$exposure[i] == 0) {
    sumu = sumu + data$outcome[i]
    countu = countu + 1
    
  } else {
    sume = sume + data$outcome[i]
    counte = counte + 1
  }
}

head(data)
tail(data)

# test proportion unexposed
sumu / countu

#test proportion exposed
sume / counte

```

## Part C

```{r c, echo = FALSE, include = FALSE}
# Part C
# cohort study
N = 50
b = 1000

datae = data[data$exposure == 1,]
datau = data[data$exposure == 0,]
estimates = data.frame(
  "seed" = rep(NA, b),
  "b0_hat" = rep(NA, b),
  "b1_hat" = rep(NA, b)
)

for (i in 1:b) {
  set.seed(i)
  exposed = sample(datae$outcome, N )
  unexposed = sample(datau$outcome, N)
  temp_dat = data.frame(
    "exposure" = c(rep(1, N), rep(0, N)),
    "outcome" = c(exposed, unexposed)
  )
  
  model = glm(outcome ~ exposure, family = binomial, data = temp_dat)
  estimates$seed[i] = i
  estimates$b0_hat[i] = summary(model)$coefficients[1, 1]
  estimates$b1_hat[i] = summary(model)$coefficients[2, 1]
}

mean_b0 = mean(estimates$b0_hat)
mean_b1 = mean(estimates$b1_hat)
median_b0 = median(estimates$b0_hat)
median_b1 = median(estimates$b1_hat)

```

```{r}
head(estimates)
tail(estimates)
```

The mean of $\beta_0$ across all simulations is `r mean_b0` and the median is `r median_b0`. The mean of $\beta_1$ across all simulations is `r mean_b1` and median is `r median_b1`. These are not quite where we would want them to be compared to part A, because we are taking a fairly small sample size ($n = 50$). As we increase the sample size, and number of iterations in the simulation, we will converge on our calculated estimates of $\beta_0$ and $\beta_1$ from part A.

## Part D

```{r d, echo = FALSE, include = FALSE}
# Part D
# case-control study
datad = data[data$outcome == 1,]
datan = data[data$outcome == 0,]
estimatesd = data.frame(
  "seed" = rep(NA, b),
  "b0_hat" = rep(NA, b),
  "b1_hat" = rep(NA, b)
)

for (i in 1:b) {
  set.seed(i)
  disease = sample(datad$exposure, N)
  nodisease = sample(datan$exposure, N)
  temp_dat = data.frame(
    "outcome" = c(rep(1, N), rep(0, N)),
    "disease" = c(disease, nodisease)
  )
  
  model = glm(outcome ~ disease, family = binomial, data = temp_dat)
  estimatesd$seed[i] = i
  estimatesd$b0_hat[i] = summary(model)$coefficients[1, 1]
  estimatesd$b1_hat[i] = summary(model)$coefficients[2, 1]  
}

dmean_b0 = mean(estimatesd$b0_hat)
dmean_b1 = mean(estimatesd$b1_hat)

dmedian_b0 = median(estimatesd$b0_hat)
dmedian_b1 = median(estimatesd$b1_hat)
```

```{r}
head(estimatesd)
tail(estimatesd)
```

The mean of $\beta_0$ across all simulations is `r dmean_b0` and the median is `r dmedian_b0`. The mean of $\beta_1$ across all simulations is `r dmean_b1` and median is `r dmedian_b1`.

## Part E

```{r e, echo = FALSE, include = FALSE}
# Part E
# cohort study
mad_b1 = mad(estimates$b1_hat)

# case-control
dmad_b1 = mad(estimatesd$b1_hat)

```

The median absolute deviance is lower for the case-control study (`r dmad_b1`) than the cohort study (`r mad_b1`). This is indicating the the efficiency of the log odds ratio may be better in the case-control style study, where we are modeling exposure based on disease, rather than modeling disease based on exposures. This is consistent with our understanding of case-control studies and cohort studies. Cohort studies require much larger populations, and case-control studies reach high power at lower population sizes.

# Appendix
```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
